Improving Item-Item Similarity Estimation during Collaborative Filtering

### 2014-03-25 paper/paper.tex
Working on Problem section

Both similarity estimates primarily leverage information from common users, that
is users who have rated both items. Because of this, they perform well with a
large number common users, but give unreliable, and dramatically varying,
results when the items have few common users. In an extreme case, if only one
user has rated item $A$ and item $B$, $PearsSim(A, B) = 1$ or $-1$. Not only is
this unlikely to be an accurate assessment of the true similarity of $A$ and
$B$, it also gives the most extreme results possible, without giving any
indication that this is a low-confidence calculation.

@@

@@

\subsection*{Technique}
While it may be impossible to get $TrueSim$ for any pair, we
can get close by calculating $Sim_n$ when a large number of users have rated
both items. Formally, we define $GoldSim$ = $Sim_N$ where $N$ is a number 
large enough that $GoldSim \approx TrueSim$.

@@

The following graph compares the results of three linear fits, with varying
degrees of fidelity to the probabilistic model. The most faithful fit uses
$\mu$, $\sigma_{s}^2$ and $\alpha$ as calculated above to predict the $GoldSim$
using Eqn (18). The next fit uses the calculated values of $\sigma_{err,n}^2$ directly,
following Eqn (16) for predictions. The third fit simply uses a series of linear
regressions, one for each $n$. This model only assumes that for each $n$,
$Sim_n$ and $GoldSim$ are linearly correlated.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{LinearTechniques.png}
	\caption{Comparison of different linear techniques. For each number of reviewers
    in common, $n$, the graphs represent the average squared error of predicted
    $GoldSim$ calculated from the observed $Sim_n$. As a baseline, the technique
    of guessing the average $GoldSim$ is included. For these experiments,
    $N=100$ for determining $GoldSim$ and $70\%$ of the data was used for
    training.}
    \label{fig:Linear}
\end{figure}

The linear regression technique trades the generality of the model for a better
fit on the training data. As the results show, the trade-off is worth it.

@@

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=1.0\textwidth]{Histograms.png}
	\caption{Histograms of {\em TrueSimR} (top) and {\em TrueSimP} (bottom) for
training datasets containing at least 40 common reviewers (left) and 80 common
reviewers (right).}
    \label{fig:Histograms}
\end{figure}

@@

\bibitem{Su2009}
    Xiaoyuan Su and Taghi M. Khoshgoftaar.
    ``A Survey of Collaborative Filtering Techniques.''
    \emph{Advances in Artificial Intelligence.} vol. 2009,
    Article ID 421425, 19 pages, 2009. doi:10.1155/2009/421425
\bibitem{Breese1998}
    Breese, John S., David Heckerman, and Carl Kadie, 
    ``Empirical analysis of predictive algorithms for collaborative filtering.'' 
    \emph{Proceedings of the Fourteenth conference on Uncertainty in artificial intelligence.}
    Morgan Kaufmann Publishers Inc., 1998.
\bibitem{Lemire1998}
    Daniel Lemire and Anna Maclachlan.
    ``Slope One Predictors for Online Rating-Based Collaborative Filtering.''
    \emph{SDM.} Vol. 5, 2005.

>>>>

\section*{Problem}
Both similarity estimates primarily leverage information from common users, that
is users who have rated both items. Because of this, they perform well with a
large number of common users, but give unreliable, and dramatically varying,
results when the items have few common users. In an extreme case, if only one
user has rated item $A$ and item $B$, $PearsSim(A, B) = 1$ or $-1$. Not only is
this unlikely to be an accurate assessment of the true similarity of $A$ and
$B$, it also gives the most extreme results possible, without giving any
indication that this is a low-confidence calculation.

@@

\subsection*{The Experiment}
We construct an experiment to test the efficacy of a predictive model. 
For either similarity metric, there is not enough data to actually calculate
$TrueSim$ for any pair. Instead, we can get close by calculating the
corresponding $Sim_n$ metric when a large number of users have rated both 
items. Formally, we define $GoldSim$ = $Sim_N$ where $N$ is a number large 
enough that $GoldSim \approx TrueSim$.

With a measurable $GoldSim$ as our predictive goal, we can implement a simple
predictivity experiment. First, we find pairs of items that have at least 
$N$ common users and calculate their $GoldSim$. Next, for each of these pairs
of items, we withhold user ratings, leaving $n$ common reviewers. This 
simulates a pair of items at times when there is a sparse amount of rating data. 
We repeat theis for $n=1,...,N$ to get varying degrees of sparsity. With the
sparse ratings alone, the model in question generates a prediction of the 
$GoldSim$, which is compared to the actual $GoldSim$. This is repeated over 
a large number of test pairs and the average squared error of the prediction 
is calculated for each $n$.

For this experimetn, we use Amazon's book data. Books are rated by users on a 1
to 5 star-scale. As in 
Some stuff about the experiment details: the data we use, how we subtract the
biases, the biasedness of the experimental dataset, training vs testing.

@@

\subsection*{Technique}
While it may be impossible to get $TrueSim$ for any pair, we
can get close by calculating $Sim_n$ when a large number of users have rated
both items. Formally, we define $GoldSim$ = $Sim_N$ where $N$ is a number 
large enough that $GoldSim \approx TrueSim$. Using this approach to approximate
$TrueSim$, we can look at the distribution of $GoldSim$ scores (Fig. 2) and see 
that the Normal distribution of $TrueSim$ is a reasonable model.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{Histograms.png}
	\caption{Histograms of {\em TrueSimR} (top) and {\em TrueSimP} (bottom) for
training datasets containing at least 40 common reviewers (left) and 80 common
reviewers (right).}
    \label{fig:Histograms}
\end{figure}

@@

The graph of Fig. 3 compares the results of three linear fits, with varying
degrees of fidelity to the probabilistic model. The most faithful fit uses
$\mu$, $\sigma_{s}^2$ and $\alpha$ as calculated above to predict the $GoldSim$
using Eqn (18). The next fit uses the calculated values of $\sigma_{err,n}^2$ directly,
following Eqn (16) for predictions. The third fit simply uses a series of linear
regressions, one for each $n$. This model only assumes that for each $n$,
$Sim_n$ and $GoldSim$ are linearly correlated. The linear regression technique 
trades the generality of the model for a better fit on the training data. 
As the results show, the trade-off is worth it.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{LinearTechniques.png}
	\caption{Comparison of different linear techniques. For each number of reviewers
    in common, $n$, the graphs represent the average squared error of predicted
    $GoldSim$ calculated from the observed $Sim_n$. As a baseline, the technique
    of guessing the average $GoldSim$ is included. For these experiments, we use
    $Sim$ = $PearSim$ and $N=100$ for determining $GoldSim$.}
    \label{fig:Linear}
\end{figure}

@@

@@

\bibitem{Su2009}
    Xiaoyuan Su and Taghi M. Khoshgoftaar.
    ``A Survey of Collaborative Filtering Techniques.''
    \emph{Advances in Artificial Intelligence.} vol. 2009,
    Article ID 421425, 19 pages, 2009. doi:10.1155/2009/421425
\bibitem{Breese1998}
    Breese, John S., David Heckerman, and Carl Kadie, 
    ``Empirical analysis of predictive algorithms for collaborative filtering.'' 
    \emph{Proceedings of the Fourteenth conference on Uncertainty in artificial intelligence.}
    Morgan Kaufmann Publishers Inc., 1998.
\bibitem{Lemire1998}
    Daniel Lemire and Anna Maclachlan.
    ``Slope One Predictors for Online Rating-Based Collaborative Filtering.''
    \emph{SDM.} Vol. 5, 2005.
\bibitem{Koren2009}
    Koren, Yehuda, Robert Bell, and Chris Volinsky. 
    ``Matrix factorization techniques for recommender systems.''
    \emph{Computer} 42.8, 2009. 30-37.
